---
title: "Data Comparison"
author: "Jeremiah Sagers"
date: "2/27/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
library(data.table)
```

```{r, include = FALSE, message=FALSE}
# Kaggle competition data
train <- read_csv("train.csv")
train <- train %>% mutate_if(is.character, as.factor)
train <- train %>% mutate_at(c('MSSubClass','MoSold'), as_factor)
test <- read_csv("test.csv")
test <- test %>% mutate_if(is.character, as.factor)
test <- test %>% mutate_at(c('MSSubClass','MoSold'), as_factor)

noprice <- rbind(train[-81], test) # combine train and test datasets

# replace NAs with actual values
for (col in names(noprice)) {
  if (is.numeric(noprice[[col]])) {
    noprice[[col]][is.na(noprice[[col]])] <- 0
  } else if (is.factor(noprice[[col]])) {
    noprice[[col]] <- as.character(noprice[[col]])
    noprice[[col]][is.na(noprice[[col]])] <- "N/A"
    noprice[[col]] <- as.factor(noprice[[col]])
  }
}

# Complete Ames housing data
house2 <- read_csv("AmesHousing.csv")
house2 <- house2 %>% mutate_if(is.character, as.factor)
house2 <- house2 %>% mutate_at(c('MS SubClass','Mo Sold'), as_factor)

house2 <- house2[-2] # remove variable not present in train/test
housenoprice <- house2[-81] # remove response variable

# make column names identical to "noprice"
names(housenoprice) <- gsub(" ", "", names(housenoprice))
names(housenoprice)[names(housenoprice) == 'Order'] <- 'Id'
names(housenoprice)[names(housenoprice) == 'YearRemod/Add'] <- 'YearRemodAdd'

for (col in names(housenoprice)) {
  if (is.numeric(housenoprice[[col]])) {
    housenoprice[[col]][is.na(housenoprice[[col]])] <- 0
  } else if (is.factor(housenoprice[[col]])) {
    housenoprice[[col]] <- as.character(housenoprice[[col]])
    housenoprice[[col]][is.na(housenoprice[[col]])] <- "N/A"
    housenoprice[[col]] <- as.factor(housenoprice[[col]])
  }
}

# Define replacement mappings
replacement_map <- c(20, 30, 40, 45, 50, 60, 70, 75, 80, 85, 90, 120, 150, 160, 180, 190)

# Replace factor levels using the mapping
housenoprice$MSSubClass <- as.factor(replacement_map[housenoprice$MSSubClass])

align_factor_levels <- function(df1, df2, column_name) {
    # Combine the factor levels from both columns
    combined_levels <- union(levels(df1[[column_name]]), levels(df2[[column_name]]))
    
    # Re-factor the columns with the combined levels
    df1[[column_name]] <- factor(df1[[column_name]], levels = combined_levels)
    df2[[column_name]] <- factor(df2[[column_name]], levels = combined_levels)
    
    return(list(df1 = df1, df2 = df2))
}

factor_columns <- names(noprice)[sapply(noprice, is.factor)]  # Get factor columns

for (col in factor_columns) {
    result <- align_factor_levels(noprice, housenoprice, col)
    noprice <- result$df1
    housenoprice <- result$df2
}
```

```{r}
# Find rows in Kaggle competition dataset that aren't in Ames dataset
df1_list <- lapply(seq_len(nrow(noprice)), function(i) noprice[i, 2:80, drop = FALSE])

matching_indices <- integer(nrow(noprice))

for (i in seq_along(df1_list)) {
    matches <- inner_join(housenoprice[2:80], df1_list[[i]], by = names(df1_list[[i]]))
    if (nrow(matches) > 0) {
        matching_indices[i] <- as.integer(rownames(matches))
    } else {
        matching_indices[i] <- 0
    }
}

which(matching_indices == 0)
```

```{r}
noprice[2525,]
filter(housenoprice, LotArea == noprice[2611,]$LotArea)

noprice[2905,which(noprice[2905,2:80] != filter(housenoprice, Id == 2904)[2:80]) + 1]
housenoprice[filter(housenoprice, Id == 2904)$Id,which(noprice[2905,2:80] != filter(housenoprice, Id == 2904)[2:80]) + 1]
```

1556: Ames has KitchenQual = Po where Kaggle has NA                             :211
1916: Ames has MSZoning = I (all) and Utilities = NoSewr instead of NA          :946
1946: Utilities = NoSewr instead of NA                                          :1014
2041: BsmtCond = Ex instead of NA                                               :1184
2152: Exterior1st, Exterior2nd = PreCast instead of NA                          :1408
2186: BsmtCond = Ex instead of NA                                               :1491
2217: MSZoning = A (agr) and Functional = Sal instead of NA                     :1554
2218: BsmtQual = Po instead of NA                                               :1555
2219: BsmtQual = Po instead of NA                                               :1557
2251: MSZoning = I (all) instead of NA                                          :1611
2421: PoolQC = TA instead of NA                                                 :1941
2474: Functional = Sal instead of NA                                            :2044
2490: SaleType = VWD instead of NA                                              :2070
2504: PoolQC = TA instead of NA                                                 :2090
2525: BsmtCond = Ex instead of NA                                               :2137
2550: MiscFeature = Elev instead of NA                                          :2181
2600: PoolQC = TA instead of NA                                                 :2279
2611: MasVnrType = CBlock instead of NA                                         :2299
2905: MSZoning = A (agr) instead of NA                                          :2904

```{r}
# Find rows in Ames that aren't in Kaggle
df_list <- lapply(seq_len(nrow(housenoprice)), function(i) housenoprice[i, 2:80, drop = FALSE])

matching_indices2 <- integer(nrow(housenoprice))

for (i in seq_along(df_list)) {
    matches <- inner_join(noprice[2:80], df_list[[i]], by = names(df_list[[i]]))
    if (nrow(matches) > 0) {
        matching_indices2[i] <- as.integer(rownames(matches))
    } else {
        matching_indices2[i] <- 0
    }
}
which(matching_indices2 == 0)
```

```{r}
housenoprice[c(107, 108, 576, 1858, 2257, 2519, 2520, 2521, 2522, 2789, 2893),]
filter(noprice, LotArea == housenoprice[c(107, 108, 576, 1858, 2257, 2519, 2520, 2521, 2522, 2789, 2893),]$LotArea)
```

107, 108, 576, 1858, 2257, 2519, 2520, 2521, 2522, 2789, and 2893 are not present in the Kaggle competition dataset.


