---
title: "EDA kaggle"
author: "Gideon Wolf"
date: "2/27/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r, include = FALSE, message=FALSE}
house <- read_csv("train.csv")
house <- house %>% mutate_if(is.character, as.factor)
house %>% mutate_at(c('MSSubClass','MoSold'), as_factor) #should we include YrSold, YrBuilt, YrRennovated, MoSold, or any of the OverallQual or OverallCond? How do we configure the year? Should we change the NAs in the Alley column to 0 or will it work with NAs?
```


# Goal: Explore the Kaggle housing dataset


## Questions: 

**How large is the data?**

**Is there a substantial amount of missing data?**

**Are there variables that vary highly with one another(covariates)?**

-Does month sold affect the price sold? 

-Is there a connection between Rennovation date and overall quality?

-Does zone and house size have any correlation?

**What distribution describes each variable?** 

**Can we make our response categorical?**

-Are there natural delineations in sales price(i.e. low, medium, high)?




```{r}

```

