---
title: "EDA kaggle"
author: "Gideon Wolf"
date: "2/27/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
library(xts)
library(ggplot2)

house <- read_csv("train.csv")
``` 


# EDA Goal: Explore the Kaggle housing dataset and prepare for model fitting

---

###Data Wrangling

We need to replace some 'NA's with another label because the data uses 'NA' as an 
indicator for some variables. The 'NA' label indicates that the house does not have 
said feature (as noted in the data_description.txt file), so we have to change 
the NA to something else to have it be considered as a level of each factor in 
the model.

```{r, message = FALSE}
NAtoNone <- c(2,7,26,31,32,33,34,36,58,59,61,64,65,73,74,75)
for (i in NAtoNone) {
  housetmp <- house
  housetmp[i][is.na(housetmp[i])] <- "None"
  house <- housetmp
}
```

Taking out the single NA value in Electrical- removing row. 
```{r}
house <- house[-1380,]
```

Now, we must change the character variables to factor variables. Additionally, 
some other variables are categories that use integer category names which must 
be changed individually.

```{r, include = FALSE, message=FALSE}
house <- house %>% mutate_if(is.character, as.factor)
house <- house %>% mutate_at('MSSubClass', as_factor) 

house$YearBuilt <- lubridate::ymd(house$YearBuilt,truncated=2L)
house$YearRemodAdd <- lubridate::ymd(house$YearRemodAdd,truncated=2L)
house$GarageYrBlt <- lubridate::ymd(house$GarageYrBlt,truncated=2L)

YrMoSold <- do.call(paste, c(sep='',list(house$YrSold),list(rep("-", times = length(house$YrSold))),list(house$MoSold),list(rep("-01", times = length(house$YrSold)))))
YrMoSold <- as_date(YrMoSold)

house <- house[-c(77,78)]
house <- as.data.frame(append(house, list(YrMoSold = YrMoSold), after = 76))
```

---


## Questions: 

**How large is the data? What is the quality of the data?**

-What type of variables are we working with?

**Is there a substantial amount of missing data?**

-If so, what kind of imputation would be most useful?

**Are there variables that vary highly with one another(covariates)?**

-Does month sold affect the price sold? 

-Is there a connection between Rennovation date and overall quality?

-Does zone and house size have any correlation?

**What distribution describes each variable?** 

**Can we make our response categorical?**

-Are there natural delineations in sales price(i.e. low, medium, high)?

-Can we predict house size based on sales price? 

***What kinds of models would be best fit for our data?***

**Once finished, how similar is the Ames dataset?**


###Size and Quality of Data

Summary and Variable Types:
```{r}
summary(house)
```

###Imputation 
Missing Data:
```{r}
md.pattern(house)

for (i in 1:length(house)) {
  print(sum(is.na(house[i])))
}
```

P: LotFrontage has 259 missing values. What imputation technique should we use to handle it?
S: MICE's PMM (predictive mean matching) can be used because we are imputing numeric data. 
```{r}
house1 = house
hosue1 = house1 %>% filter(LotFrontage < 200) ## This doesn't work

ggplot(data = house) +
  geom_boxplot(aes(x = MSSubClass, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = MSZoning, y = LotFrontage))

ggplot(data = house,aes(x = LotArea, y = LotFrontage)) +
  geom_point() +
  geom_smooth()

ggplot(data = house) +
  geom_boxplot(aes(x = Street, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = Alley, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LotShape, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LandContour, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LotConfig, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LandSlope, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = Condition1, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = Condition2, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = BldgType, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = HouseStyle, y = LotFrontage))

ggplot(data = house,aes(x = YearBuilt, y = LotFrontage)) +
  geom_point() +
  geom_smooth()

ggplot(data = house) +
  geom_boxplot(aes(x = Foundation, y = LotFrontage))

ggplot(data = house1 ,aes(x = GrLivArea, y = LotFrontage)) +
  geom_point() +
  geom_smooth()

cor(house$LotFrontage, house$LotArea) ## Returns NA
```
```{r}
numhouse <- select_if(house, is.numeric)
imputed_house <- mice(numhouse, 
                       m = 10,  
                       method = 'pmm', 
                       seed = 1)
imputed_house$loggedEvents
```

P: While this command runs, many of the variables are marked as "(nearly) multi-collinear." The logged events command asks "Does your dataset contain duplicates, linear transformation, or factors with unique respondent names?" How do we approach this problem?
S: Since it is just a warning we can leave it for now and return to it later if we have time. 

```{r}
set.seed(1)
imphouse <- complete(imputed_house)
housetmp <- house

#### Repair this, takes out many columns/values for no reason
#for (i in names(imphouse)[2:length(names(imphouse))]) {
#  housetmp[i] <- imphouse[match(housetmp$ID, imphouse$ID)]
#}
####


house <- housetmp
summary(house)
md.pattern(house)
```
Now we only have one column with NA's left. 

### Handling GarageYrBuilt
---
P: This variable is tricky. It is the date of garage construction with an NA meaning that there is no garage attached to the house. To impute an NA for this variable would be meaningless because there would be no garage to have been built on an imputed day- imputation would ignore the real-world meaning of this variable and potentially skew sales price. However, leaving the NA's without imputation could impede upon model creation. Additionally, this variable could hold important information about the sales price, so it may be unwise to simply take the column out without knowing if it is valuable. 
S: We will investigate the similarity between other garage variables and GarageYrBlt to see if the impact of GarageYrBlt on SalePrice can be summarized by other variables. If other variables cannot summarize what GarageYrBlt can provide to models, we must create a categorical variable for GarageYrBlt. We will balance the distribution of the variable along with real-world phenomena that may have had an impact on SalePrice

If we are to include GarageYrBlt, how do we handle the data if there is no garage? If we change it to a factor, what categories do we split them into? (old, mid, new OR 1900-1910, ..., 2000-2010) And should the categories be chosen so each category has an even numbers of observations or maintain the distribution of observations in the categories? Should we use the quartiles to choose the delineations? 
If we use imputeTS package, should we remove other date columns? 
```{r}
hist(house$GarageYrBlt,breaks = 30)
ggplot(aes(x = GarageYrBlt, y = GarageArea, color = SalePrice), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = GarageQual), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = GarageCond), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = GarageFinish), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = GarageType), data = house) +
  geom_point()
```



---

### Other Q?

How does the 2008 housing crash affect prices, does it at all? Can we use this pattern to predict other crashes in history?

How does sale price vary over time? 

What years do we have the most data? 

How does the relationship between sale price and number of houses over time covary? 
```{r}
hist(house$YrMoSold,breaks = 30)
plot(house$YrMoSold,house$SalePrice)
```




