---
title: "EDA Kaggle"
author: "Gideon Wolf"
date: "2/27/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
library(xts)
library(ggplot2)

house <- read_csv("train.csv")
``` 


# EDA Goal: Explore the Kaggle housing dataset and prepare for model fitting

---

## Data Handling {.tabset}

### Data Wrangling
We need to replace some 'NA's with another label because the data uses 'NA' as an 
indicator for some variables. The 'NA' label indicates that the house does not have 
said feature (as noted in the data_description.txt file), so we have to change 
the NA to something else to have it be considered as a level of each factor in 
the model.

```{r, message = FALSE}
NAtoNone <- c(2,7,26,31,32,33,34,36,58,59,61,64,65,73,74,75)
for (i in NAtoNone) {
  housetmp <- house
  housetmp[i][is.na(housetmp[i])] <- "None"
  house <- housetmp
}
```

Taking out the single NA value in Electrical- removing row. 
```{r}
house <- house[-1380,]
```

Now, we must change the character variables to factor variables. Additionally, 
some other variables are categories that use integer category names which must 
be changed individually.

```{r, include = FALSE, message=FALSE}
house <- house %>% mutate_if(is.character, as.factor)
house <- house %>% mutate_at('MSSubClass', as_factor) 

YrMoSold <- do.call(paste, c(sep='',list(house$YrSold),list(rep("-", times = length(house$YrSold))),list(house$MoSold),list(rep("-01", times = length(house$YrSold)))))
YrMoSold <- as_date(YrMoSold)
house <- as.data.frame(append(house, list(YrMoSold = YrMoSold), after = 78))
```


### Size and Quality of Data

Summary and Variable Types:

```{r}
summary(house)
```

**ADD**: what does this summary tell us?


### Handling GarageYrBuilt

P: This variable is tricky. It is the date of garage construction with an NA meaning that there is no garage attached to the house. To impute an NA for this variable would be meaningless because there would be no garage to have been built on an imputed day- imputation would ignore the real-world meaning of this variable and potentially skew sales price. However, leaving the NA's without imputation could impede upon model creation. Additionally, this variable could hold important information about the sales price, so it may be unwise to simply take the column out without knowing if it is valuable. 

S: We will investigate the similarity between other garage variables and GarageYrBlt to see if the impact of GarageYrBlt on SalePrice can be summarized by other variables. After analyzing the relationship between GarageYrBlt, Sales Price, and other garage variables, we feel comfortable taking out the variable entirely. We will also try and create a categorical variable for GarageYrBlt (1900-1940 old, 1940-1980 mid, 1980-2010 new). 

```{r}
p <- print(select(house, row.names = c("GarageYrBlt", "GarageArea")))
p[p[2]== 0,]
p[is.na(p[1]),] 
```

The above shows that each instance of an NA date corresponds to a 0 GarageArea and vice versa. 

```{r}
hist(house$GarageYrBlt,breaks = 30)
ggplot(aes(x = GarageYrBlt, y = SalePrice), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = SalePrice, color = GarageArea), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = SalePrice, color = GarageQual), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = SalePrice, color = GarageCond), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = SalePrice, color = GarageFinish), data = house) +
  geom_point()
ggplot(aes(x = GarageYrBlt, y = SalePrice, color = GarageType), data = house) +
  geom_point()
```

**ADD**: How do we interpret these graphs?

```{r}
house <- house[-60] # Taking out GarageYrBlt
```


### Imputation 

Missing Data:

```{r}
md.pattern(house)

for (i in 1:length(house)) {
  print(sum(is.na(house[i])))
}
```

P: LotFrontage has 259 missing values. What imputation technique should we use to handle it?

S: MICE's PMM (predictive mean matching) can be used because we are imputing numeric data. 

```{r}
house1 = house
hosue1 = house1 %>% filter(LotFrontage < 200) ## This doesn't work

ggplot(data = house) +
  geom_boxplot(aes(x = MSSubClass, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = MSZoning, y = LotFrontage))

ggplot(data = house,aes(x = LotArea, y = LotFrontage)) +
  geom_point() +
  geom_smooth()

ggplot(data = house) +
  geom_boxplot(aes(x = Street, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = Alley, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LotShape, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LandContour, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LotConfig, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = LandSlope, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = Condition1, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = Condition2, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = BldgType, y = LotFrontage))

ggplot(data = house) +
  geom_boxplot(aes(x = HouseStyle, y = LotFrontage))

ggplot(data = house,aes(x = YearBuilt, y = LotFrontage)) +
  geom_point() +
  geom_smooth()

ggplot(data = house) +
  geom_boxplot(aes(x = Foundation, y = LotFrontage))

ggplot(data = house1 ,aes(x = GrLivArea, y = LotFrontage)) +
  geom_point() +
  geom_smooth()

cor(house$LotFrontage, house$LotArea) ## Returns NA
```

**ADD**: description of these graphs, what does it mean that the cov function returns an NA?

```{r}
numhouse <- select_if(house, is.numeric)
imputed_house <- mice(numhouse, 
                       m = 10,  
                       method = 'pmm', 
                       seed = 1)
imputed_house$loggedEvents
```

P: While this command runs, many of the variables are marked as "(nearly) multi-collinear." The logged events command asks "Does your dataset contain duplicates, linear transformation, or factors with unique respondent names?" How do we approach this problem?

S: Since it is just a warning we can leave it for now and return to it later if we have time. 

```{r}
set.seed(1)
imphouse <- complete(imputed_house)
housetmp <- house
for (i in names(imphouse)[2:length(names(imphouse))]) {
  housetmp[i] <- imphouse[i]
}
house <- housetmp
summary(house)
md.pattern(house)
```
Now we only have one column with NA's left: GarageYrBlt. 



### Outliers

```{r}
housetmp <- select_if(house, is.numeric)
for (i in 1:length(housetmp)) {
  print(names(housetmp[i]))
  col <- housetmp[,i]
  z <- vector("numeric", length = length(col))
  outliers <- vector("numeric")
  for (j in 1:length(col)) {
    z[j] <- (col[j]-mean(col))/sd(col)
    outliers <- append(outliers, if(abs(z[j]) > 5) j) }
  cat("Index of outliers: ", outliers, "\n")
  cat("Value of outliers: ", col[outliers], "\n")
  cat("           Min, 1st, Med, Mean, 3rd, Max\n")
  cat("Quartiles: ", summary(housetmp[,i]), "\n\n")
}
```

**ADD**: What to do with these outliers? Some of these variables are categorical variables so we will probably not take those "outliers" out. It is currently choosing outliers 5 standard deviations away from the mean, is that fine or should it be changed? 

---

## Questions: 

All answered above: 
**How large is the data? What is the quality of the data?**
-What type of variables are we working with?
**Is there a substantial amount of missing data?**
-If so, what kind of imputation would be most useful?
-Do we have any outliers?

To answer: 
**Are there variables that vary highly with one another(covariates)?**

-Does month sold affect the price sold? 
```{r}
ggplot(data = house, aes(x = YrMoSold, y = SalePrice)) +
  geom_point() +
  geom_smooth(method='lm',se = TRUE)

h_lm <- lm(SalePrice~YrMoSold, data= house)
plot(h_lm)
```
Slope ~= 0, no relationship

-Is there a connection between Rennovation date and overall quality?
```{r}
ggplot(data = house, aes(x = YearRemodAdd, y = OverallCond)) +
  geom_jitter() +
  geom_smooth(method='lm',se = TRUE)

h_lm <- lm(SalePrice~YrMoSold, data= house)
plot(h_lm)
```
slope ~= 0, no correlation relationship.

-Does zone and house size have any correlation?
```{r}

```

**What distribution describes each variable?** 

**Can we make our response categorical?**

-Are there natural delineations in sales price(i.e. low, medium, high)?

-Can we predict house size based on sales price? 

***What kinds of models would be best fit for our data?***
lazypredict python

**Once finished, how similar is the Ames dataset?**

-How does the 2008 housing crash affect prices, does it at all?

-How does sale price vary over time? 

-What years do we have the most data? 

-How do sale price and number of houses in the data per year relate? 

```{r}
hist(house$YrMoSold,breaks = 30)
plot(house$YrMoSold,house$SalePrice)
```

**ADD**: description of these graphs


## Modelling Stuff

"lazy()" function in R (alongside predict.lazy())
