---
title: "EDA kaggle"
author: "Gideon Wolf"
date: "2/27/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
library(xts)
library(ggplot2)

house <- read_csv("train.csv")
``` 

We need to replace some 'NA's with another label because the data uses 'NA' as an 
indicator for some variables. The 'NA' label indicates that the house does not have 
said feature (as noted in the data_description.txt file), so we have to change 
the NA to something else to have it be considered as a level of each factor in 
the model.

```{r, message = FALSE}
NAtoNone <- c(2,7,26,31,32,33,34,36,58,59,61,64,65,73,74,75)
for (i in NAtoNone) {
  housetmp <- house
  housetmp[i][is.na(housetmp[i])] <- "None"
  house <- housetmp
}
```

Now, we must change the character variables to factor variables. Additionally, 
some other variables are categories that use integer category names which must 
be changed individually.

```{r, include = FALSE, message=FALSE}
house <- house %>% mutate_if(is.character, as.factor)
house <- house %>% mutate_at('MSSubClass', as_factor) 

house$YearBuilt <- lubridate::ymd(house$YearBuilt,truncated=2L)
house$YearRemodAdd <- lubridate::ymd(house$YearRemodAdd,truncated=2L)
house$GarageYrBlt <- lubridate::ymd(house$GarageYrBlt,truncated=2L)

YrMoSold <- do.call(paste, c(sep='',list(house$YrSold),list(rep("-", times = length(house$YrSold))),list(house$MoSold),list(rep("-01", times = length(house$YrSold)))))
YrMoSold <- as_date(YrMoSold)

house <- house[-c(77,78)]
house <- as.data.frame(append(house, list(YrMoSold = YrMoSold), after = 76))
```


# Goal: Explore the Kaggle housing dataset


## Questions: 

**How large is the data? What is the quality of the data?**

-What type of variables are we working with?

**Is there a substantial amount of missing data?**

-If so, what kind of imputation would be most useful?

**Are there variables that vary highly with one another(covariates)?**

-Does month sold affect the price sold? 

-Is there a connection between Rennovation date and overall quality?

-Does zone and house size have any correlation?

**What distribution describes each variable?** 

**Can we make our response categorical?**

-Are there natural delineations in sales price(i.e. low, medium, high)?

-Can we predict house size based on sales price? 

***What kinds of models would be best fit for our data?***

**Once finished, how similar is the Ames dataset?**


#Size and Quality of Data

Summary and Variable Types:
```{r}
summary(house)
```


Missing Data:
```{r}
md.pattern(house)

for (i in 1:length(house)) {
  print(sum(is.na(house[i])))
}
```

LotFrontage has 259 missing values. What imputation technique should we use to handle it?

If we are to include GarageYrBlt, how do we handle the data if there is no garage? If we change it to a factor, what categories do we split them into? (old, mid, new OR 1900-1910, ..., 2000-2010) And should the categories be chosen so each category has an even numbers of observations or maintain the distribution of observations in the categories? Should we use the quartiles to choose the delineations? 
```{r}
hist(house$GarageYrBlt,breaks = 30)
```



How does the 2008 housing crash affect prices, does it at all? Can we use this pattern to predict other crashes in history?

How does sale price vary over time? 

What years do we have the most data? 

How does the relationship between sale price and number of houses over time covary? 
```{r}
hist(house$YrMoSold,breaks = 30)
plot(house$YrMoSold,house$SalePrice)
```




