---
title: "EDA kaggle"
author: "Gideon Wolf"
date: "2/27/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)

house <- read_csv("train.csv")
```


```{r, message = FALSE}
NAtoNone <- c(7,26,31,32,33,34,36,58,59,61,64,65,73,74,75)
for (i in NAtoNone) {
  housetmp <- house
  #housetmp[i] <- as.character(housetmp[i])
  housetmp[i][is.na(housetmp[i])] <- "None"
  #housetmp$Alley <- as.factor(housetmp$Alley)
  house <- housetmp
}
```

We need to do this with these variables because each one has NA as an 
indicator that the house does not have that feature (as noted in the 
data_description.txt file), so we have to change the NA to something else to 
have it be considered as a level of each factor in the model.

```{r, include = FALSE, message=FALSE}
house <- house %>% mutate_if(is.character, as.factor)
house <- house %>% mutate_at(c('MSSubClass','MoSold'), as_factor) 
#should we include YrSold, YrBuilt, YrRennovated, 
# MoSold, or any of the OverallQual or OverallCond? How do
#we configure the year? Should we change the NAs in the
#Alley column to 0 or will it work with NAs?
```


# Goal: Explore the Kaggle housing dataset


## Questions: 

**How large is the data? What is the quality of the data?**

-What type of variables are we working with?

**Is there a substantial amount of missing data?**

-If so, what kind of imputation would be most useful?

**Are there variables that vary highly with one another(covariates)?**

-Does month sold affect the price sold? 

-Is there a connection between Rennovation date and overall quality?

-Does zone and house size have any correlation?

**What distribution describes each variable?** 

**Can we make our response categorical?**

-Are there natural delineations in sales price(i.e. low, medium, high)?

-Can we predict house size based on sales price? 

***What kinds of models would be best fit for our data?***

**Once finished, how similar is the Ames dataset?**


#Size and Quality of Data

Summary and Variable Types:
```{r}
summary(house)
```


Missing Data:
```{r}
md.pattern(house)
```

