---
title: "EDA kaggle"
author: "Gideon Wolf"
date: "2/27/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
```

```{r, include = FALSE, message=FALSE}
house <- read_csv("train.csv")
house <- house %>% mutate_if(is.character, as.factor)
house <- house %>% mutate_at(c('MSSubClass','MoSold'), as_factor) #should we include YrSold, YrBuilt, YrRennovated, 
# MoSold, or any of the OverallQual or OverallCond? How do
#we configure the year? Should we change the NAs in the
#Alley column to 0 or will it work with NAs?
```

```{r}
NAtoNone <- c("")
housetmp <- house
housetmp$Alley <- as.character(housetmp$Alley)
housetmp[7][is.na(housetmp[7])] <- "None"
housetmp$Alley <- as.factor(housetmp$Alley)
house <- housetmp
```

We need to do ^ with all of the following variables (each one has NA as an 
indicator that the house does not have that feature, so we have to change the NA
to something else to have it be considered in the model): 
Fence
PoolQC
FireplaceQu
MiscFeature
GarageCond
GarageQual
GarageFinish
GarageYrBlt
GarageType
BsmtFinType2 
BsmtFinType1 
BsmtExposure 
BsmtCond 
BsmtQual 
How do you 



# Goal: Explore the Kaggle housing dataset


## Questions: 

**How large is the data? What is the quality of the data?**

-What type of variables are we working with?

**Is there a substantial amount of missing data?**

-If so, what kind of imputation would be most useful?

**Are there variables that vary highly with one another(covariates)?**

-Does month sold affect the price sold? 

-Is there a connection between Rennovation date and overall quality?

-Does zone and house size have any correlation?

**What distribution describes each variable?** 

**Can we make our response categorical?**

-Are there natural delineations in sales price(i.e. low, medium, high)?

-Can we predict house size based on sales price? 

***What kinds of models would be best fit for our data?***

**Once finished, how similar is the Ames dataset?**


#Size and Quality of Data

Summary and Variable Types:
```{r}
summary(house)
```


Missing Data:
```{r}
md.pattern(house)
```

